---
title: "US Real Estate Market Presentation"
date: today
author: 
  - Emir Hamulic
  - Elias Gabriel
self-contained: true
code-link: true
code-tools: true
lightbox: true  ## click on image to open in lightbox
format: 
  revealjs:
    scrollable: true
    smaller: true
    theme: sky ##https://quarto.org/docs/presentations/revealjs/themes.html
---

```{r}
#| code-summary: Libraries
library <- function(...) {suppressPackageStartupMessages(base::library(...))}
library(tidyverse)
library(dplyr)
library(knitr)
library(tidyr)
library(rmarkdown)
library(janitor)
library(scales)
library(tidytext)
library(ggforce)
library(GGally)
library(DT)
library(kableExtra)
library(broom)
library(plotly)
```

```{r}
data = read.csv("data/realtor-data.zip.csv") # Data import
data = subset(data, select = c(status, price, bed, bath, acre_lot, city, state, house_size)) # keep relevant columns
data$status = as.factor(data$status)
data$city = as.factor(data$city)
data$state = as.factor(data$state)
data <- na.omit(data)
data = data |>
  filter(price > 10000 & price < 1000000000)
data = data |>
  mutate(price_per_sqm = price/house_size)
```

## Basics
- Dataset of 2.2M+ real estate listings across the United States (source: Kaggle, Ahmed Shahriar Sakib).
- Includes key features: price, size, lot area, bedrooms, bathrooms, location, and status.
- Goals:
  - Understand price patterns and housing characteristics across states.
  - Analyze relationships between property features and regional affordability trends.
  - Provide insights into the current U.S. housing market dynamics.

## Data structure
```{r}
#| label: "data_structure_after_transformation"
#| tbl-cap: "Structure of cleaned dataset"
#| code-fold: true

structure_tbl <- tibble::tibble(
  Variable = names(data),
  Type = sapply(data, function(x) class(x)[1]),
  Example = sapply(data, function(x) {
    val <- unique(x[!is.na(x)])[1]
    if (is.factor(val)) as.character(val) else as.character(round(val, 2))
  }),
  Missing = sapply(data, function(x) sum(is.na(x)))
)

kable(
  structure_tbl,
  caption = "Structure summary of the dataset",
  align = c("l", "l", "l", "r")
)
```

## Distribution of Numerical Variables

```{r}
#| label: "Logarithmic Visualisation"
#| tbl-cap: "Visualisation of numerical variables in dataframe"
#| code-fold: true
data |>
  clean_names() |>
  pivot_longer(cols = where(is.numeric)) |>
  ggplot(aes(x = value, fill = name)) +
  geom_histogram(bins = 30, alpha = 0.7, color = "white") +
  scale_x_log10(labels = label_comma()) +   # 👈 echte Werte, log-Skala
  facet_wrap(~ name, scales = "free_x") +
  theme_minimal() +
  labs(
    title = "Distribution of Numerical Variables (logarithmic scale)",
    x = "Value",
    y = "Count"
  ) +
  theme(
    legend.position = "none",
    axis.text.x = element_text(angle = 25, hjust = 1) 
    )
```

## Top Categories per Variable

```{r}
#| label: "Visualization Nominal Statistics"
#| tbl-cap: "Top categories for factor variables with counts, proportions, and mean price"
#| code-fold: true

top_n_per_var <- 10 

nominal_summary <- data |>
  clean_names() |>
  select(where(is.factor), price) |>
  pivot_longer(cols = where(is.factor),
               names_to = "Variable",
               values_to = "Category") |>
  group_by(Variable, Category) |>
  summarise(
    Count = n(),
    Percent = round(100 * Count / nrow(data), 2),
    Mean_Price = round(mean(price, na.rm = TRUE), 0),
    .groups = "drop"
  ) |>
  group_by(Variable) |>
  slice_max(order_by = Count, n = top_n_per_var, with_ties = FALSE) |>
  ungroup()


nominal_summary <- nominal_summary |>
  group_by(Variable) |>
  mutate(Category = forcats::fct_reorder(Category, Count),
         Category = factor(Category, levels = unique(Category))) |>
  ungroup()

# Plot: Facets untereinander, mit eigener x-Skala und y-Skala pro Variable
ggplot(nominal_summary, aes(x = Count, y = Category, fill = Variable)) +
  geom_col(show.legend = FALSE, alpha = 0.8, width = 0.7) +
  facet_wrap(~ Variable, ncol = 1, scales = "free", drop = TRUE) +
  scale_x_continuous(labels = label_comma()) +   #Tausendertrennung, keine 1e+05
  theme_minimal() +
  labs(
    title = "Top Categories per Factor Variable",
    x = "Count",
    y = "Category"
  ) +
  theme(
    panel.spacing.y = unit(1, "lines"),
    strip.text = element_text(size = 12, face = "bold"),
    axis.text.y = element_text(size = 8),
    plot.margin = margin(5, 15, 5, 5)
  )
```

## Pairs Plot
```{r}
#| label: "Pairs Plot"
#| code-fold: true

set.seed(123)

data_num <- data |>
  janitor::clean_names() |>
  select(where(is.numeric)) |>
  slice_sample(n = 3000) |>
  mutate(across(everything(), log1p)) 

p <- ggpairs(
  data_num,
  progress = FALSE,
  upper = list(continuous = wrap("cor", size = 4, alignPercent = 0.8, stars = TRUE)),
  lower = list(continuous = wrap("points", alpha = 0.3, size = 0.7)),
  diag = list(continuous = wrap("densityDiag", alpha = 0.7))
)

p +
  theme_minimal(base_size = 11) +
  theme(
    strip.text = element_text(size = 8, face = "bold"),
    panel.grid = element_blank(),
    axis.text = element_text(size = 8),
    axis.title = element_text(size = 9),
    plot.title = element_text(face = "bold", size = 14, hjust = 0.5)
  ) +
  labs(title = "Pairs Plot (log-transformiert, n=3000)")
```

## Price vs House Size
```{r}
#| label: "scatter_price_size"
#| tbl-cap: "Price vs. house size by listing status"
#| code-fold: true

# Stichprobe ziehen für Performance
set.seed(123)
sample_data <- data %>% sample_n(50000)

plot_ly(
  sample_data,
  x = ~house_size,
  y = ~price,
  color = ~status,
  type = "scatter",
  mode = "markers",
  alpha = 0.6
) %>%
  plotly::layout(
    title = list(text = "Relationship Between House Size and Price by Status"),
    xaxis = list(title = "House Size (sqft)"),
    yaxis = list(title = "Price ($)", type = "log")
  )
```

## Regression
```{r}
#| code-fold: true

model <- lm(price ~ house_size + bath + bed + state + status, data = data)

tidy(model, conf.int = TRUE) |>
  filter(term != "(Intercept)") |>
  mutate(term = reorder(term, estimate)) |>
  ggplot(aes(x = estimate, y = term, fill = estimate > 0)) +
  geom_col(show.legend = FALSE) +
  geom_vline(xintercept = 0, linetype = "dashed") +
  theme_minimal() +
  labs(
    title = "Greatest contributors to price",
    x = "regression coefficient",
    y = ""
  )
```

## Average Price Map

```{r}
#| label: "map_avg_price"
#| tbl-cap: "Average property price by U.S. state"
#| code-fold: true

valid_states <- tibble(
  state_name = c(state.name, "District of Columbia"),
  state_abbr = c(state.abb,  "DC")
)

map_price <- data |>
  group_by(state) |>
  summarise(avg_price = mean(price, na.rm = TRUE), .groups = "drop") |>
  inner_join(valid_states, by = c("state" = "state_name")) |>
  mutate(avg_price_k = avg_price / 1000)

plot_ly(
  map_price,
  type = "choropleth",
  locationmode = "USA-states",
  locations = ~state_abbr,
  z = ~avg_price_k,
  text = ~paste0(state, "<br>Avg Price: $", round(avg_price_k, 1), "K"),
  colorscale = list(c(0, 1), c("lightblue", "darkblue")),
  colorbar = list(title = "Avg Price ($K)")
) |>
  plotly::layout(
    title = list(text = "Average Property Price by U.S. State"),
    geo = list(scope = "usa", projection = list(type = "albers usa"))
  )
```

## Average Size Map

```{r}
#| label: "map_avg_size"
#| tbl-cap: "Average house size by U.S. state"
#| code-fold: true

map_size <- data %>%
  group_by(state) %>%
  summarise(avg_size = mean(house_size, na.rm = TRUE), .groups = "drop") %>%
  inner_join(valid_states, by = c("state" = "state_name"))

plot_ly(
  map_size,
  type = "choropleth",
  locationmode = "USA-states",
  locations = ~state_abbr,
  z = ~avg_size,
  text = ~paste0(state, "<br>Avg Size: ", round(avg_size), " sqft"),
  colorscale = list(c(0, 1), c("lightgreen", "darkgreen")),
  colorbar = list(title = "Avg Size (sqft)")
) %>%
  plotly::layout(
    title = list(text = "Average House Size by U.S. State"),
    geo = list(scope = "usa", projection = list(type = "albers usa"))
  )
```

## Price Range Map
```{r}
#| label: "map_extreme_price"
#| tbl-cap: "Price range (max − min) by U.S. state"
#| code-fold: true

map_extremes <- data |>
  group_by(state) |>
  summarise(
    min_price = suppressWarnings(min(price, na.rm = TRUE)),
    max_price = suppressWarnings(max(price, na.rm = TRUE)),
    .groups = "drop"
  ) |>
  mutate(range_price = max_price - min_price) |>
  inner_join(valid_states, by = c("state" = "state_name"))

plot_ly(
  map_extremes,
  type = "choropleth",
  locationmode = "USA-states",
  locations = ~state_abbr,
  z = ~range_price,
  text = ~paste0(
    state,
    "<br>Min: $", formatC(min_price, big.mark = ",", format = "f", digits = 0),
    "<br>Max: $", formatC(max_price, big.mark = ",", format = "f", digits = 0)
  ),
  colorscale = "Reds",
  colorbar = list(title = "Price Range ($)")
) |>
  plotly::layout(
    title = list(text = "Price Extrem Values by U.S. State (Max − Min)"),
    geo = list(scope = "usa", projection = list(type = "albers usa"))
  )
```

## Thank you for your attention
