---
title: "Data Exploration Gabriel Hamulic"
subtitle: "Dataset: US Real Estate Dataset"
author: "Hamulic, Gabriel"
date: today
embed-resources: true
format:
  html:
    output-file: US-Real-Estate_Gabriel_Hamulic.html
    #output-ext: "html.html" 
    toc: true
    toc-location: right
    code-link: true
    code-tools: true
    #df-print: kable
    theme:
      light: flatly
      dark: darkly
    #echo: fenced
  pdf: 
    output-file: US-Real-Estate_Gabriel_Hamulic.pdf 
    toc: true
    number-sections: true
    code-link: true
    df-print: tibble
    crossref: 
      lof-title: "List of Figures"
fig-align: center
execute: 
  warning: false
---

\listoffigures 
\listoftables
\listoflistings

::: callout-caution
### Instructions

Your report must be of high quality, meaning that your report:

-   is visually and textually pleasing of
-   does not look/read/feel like a draft instead of a finished analysis
-   explains/discusses your findings and results in the main text, e.g., explain/discuss all figures/table in the main text
-   is representable such that it can show to any interested third party
-   uses figure/table captions/linking/reference (see example further down)
-   Do not show any standard printout of R-code, use for data.frame/tibbles `knitr::kable()` printing.
-   Do not simply print datasets (too many lines) use instead `rmarkdown::paged_table()`
:::

{{< pagebreak >}}

# Introduction

In this project, we explore a large dataset of over 2 million real estate listings across the United States. The data includes key property attributes such as price, size, lot area, location, and listing status.

Our goal is to gain a deeper understanding of the current U.S. housing market by identifying:

- Patterns in property prices and sizes across states and cities

- Relationships between housing characteristics (e.g., size, bedrooms, bathrooms)

- Regional differences in affordability and market activity

## Libraries

```{r}
#| code-summary: Libraries
#| code-fold: true
library <- function(...) {suppressPackageStartupMessages(base::library(...))}
library(tidyverse)
library(dplyr)
library(knitr)
library(tidyr)
library(rmarkdown)
library(janitor)
library(scales)
library(tidytext)
library(ggforce)
library(GGally)
library(DT)
library(kableExtra)
library(broom)
library(plotly)
```

# Data

## Data source

In this data exploration we are looking at the US Real Estate market with use of a dataset from kaggle published by Ahmed Shahriar Sakib. It contains over 2.2 Million Real Estate listings broken down to State, Size, Price (among other factors). (Source: <https://www.kaggle.com/datasets/ahmedshahriarsakib/usa-real-estate-dataset/data>)

## Data import

```{r setup, include=FALSE}
options(dplyr.print_max = 15, dplyr.print_min = 10)
```

```{r}
data = read.csv("data/realtor-data.zip.csv") # Data import
```

## Data Transformation

```{r}
data = subset(data, select = c(status, price, bed, bath, acre_lot, city, state, house_size)) # keep relevant columns
```

### Structure with standard datatypes
```{r}
#| label: "data_structure"
#| tbl-cap: "Structure of cleaned dataset"
#| code-fold: true

structure_tbl <- tibble::tibble(
  Variable = names(data),
  Type = sapply(data, function(x) class(x)[1]),
  Example = sapply(data, function(x) {
    val <- unique(x[!is.na(x)])[1]
    if (is.factor(val)) 
      as.character(val)
  }),
  Missing = sapply(data, function(x) sum(is.na(x)))
)

kable(
  structure_tbl,
  caption = "Structure summary of the dataset",
  align = c("l", "l", "l", "r")
)
```

```{r}
#| code-fold: true

# Assign Data Types
data$status = as.factor(data$status)
data$city = as.factor(data$city)
data$state = as.factor(data$state)
```

### NA Removal
```{r}
#| code-fold: true
before_rows <- nrow(data)
data <- na.omit(data)
after_rows <- nrow(data)
kable(data.frame(Description = c("Before NA removal", "After NA removal"),
Rows = c(before_rows, after_rows)))
```

The dataset now has `r nrow(data)` observations and `r ncol(data)` variables after removing rows with missing values.

### Filtering
```{r}
#| code-fold: true

# Filter min and max values
data = data |>
  filter(price > 10000 & price < 1000000000)
```

### Calculations
```{r}
#| code-fold: true

data = data |>
  mutate(price_per_sqm = price/house_size)
```

### Cleaned Dataset
```{r}
#| code-fold: true

paged_table(data)
```

### Structure after transformation

```{r}
#| label: "data_structure_after_transformation"
#| tbl-cap: "Structure of cleaned dataset"
#| code-fold: true

structure_tbl <- tibble::tibble(
  Variable = names(data),
  Type = sapply(data, function(x) class(x)[1]),
  Example = sapply(data, function(x) {
    val <- unique(x[!is.na(x)])[1]
    if (is.factor(val)) as.character(val) else as.character(round(val, 2))
  }),
  Missing = sapply(data, function(x) sum(is.na(x)))
)

kable(
  structure_tbl,
  caption = "Structure summary of the dataset",
  align = c("l", "l", "l", "r")
)
```


After cleaning the dataset, all variables have appropriate data types and no missing values (**n = `r nrow(data)`**):

- **status** – Factor variable showing the listing status (e.g., for_sale, sold).

- **price** – Numeric value for the property’s price in USD.

- **bed, bath** – Integer counts of bedrooms and bathrooms.

- **acre_lot** – Numeric size of the lot (in acres).

- **city, state** – Factor variables identifying the property’s location.

- **house_size** – Numeric size of the house (in square feet).

- **price_per_sqm** – Numeric variable derived from price / house_size to compare prices across properties.

All rows with missing data were removed, and categorical variables were converted to factors for easier analysis and visualization later on.

## Data dictionary

```{r}
#| code-fold: true

tibble(
  Variable = c("price", "status", "acre_lot", "state", "house_size"),
  Description = c(
    "The price for which the item was listed on the market",
    "The status if the house is already sold or still for sale",
    "The size of the land / lot on which the house is located in acres",
    "The state in which the house is located",
    "The size of the house in square feet"
  )
) |>
  kable(
    caption = "Description of key variables in the dataset",
    align = c("l", "l")
  )
```


# Summary statistic tables

In this section we will cover the summary of our cleaned dataset. We will explore basic statistical values from our data.

## Numeric Statistics

### Summary of numerical values

```{r}
#| label: "Numeric Statistics"
#| tbl-cap: "Summary statistics of numerical variables in dataframe"
#| code-fold: true
data |> 
  janitor::clean_names() |>
  mutate(row = row_number() |> factor()) |> 
  pivot_longer(cols = where(is.numeric)) |> 
  group_by(name) |> 
  summarize(N = n(),
            min = min(value),
            mean = mean(value),
            median = median(value),
            max = max(value),
            st.dev = sd(value)
            ) |> 
  knitr::kable(digits = 2)
```

#### Interpretation

**price**: Very wide range (\$10.4k–\$51.5M). Mean (\$573k),  median (\$379k), indicating strong right-skew and high-priced outliers.

**house_size**: Average ~2,119 sqft, median 1,812 sqft. Extremely large max (1,560,780 sqft) signals outliers. The distribution is right-skewed.

**acre_lot**: Median 0.21 acres vs. mean 12.75 acres → a few very large parcels inflate the mean.

**bed / bath**: Typical homes (~3 beds, 2 baths) with modest spread; minima at 1 suggest realistic counts.

**price_per_sqm**: Mean \$262.42 vs. median \$197.42, also right-skewed, consistent with price outliers.

### Visualisation of numerical values

```{r}
#| label: "Logarithmic Visualisation"
#| tbl-cap: "Visualisation of numerical variables in dataframe"
#| code-fold: true
data |>
  clean_names() |>
  pivot_longer(cols = where(is.numeric)) |>
  ggplot(aes(x = value, fill = name)) +
  geom_histogram(bins = 30, alpha = 0.7, color = "white") +
  scale_x_log10(labels = label_comma()) +   # 👈 echte Werte, log-Skala
  facet_wrap(~ name, scales = "free_x") +
  theme_minimal() +
  labs(
    title = "Distribution of Numerical Variables (logarithmic scale)",
    x = "Value",
    y = "Count"
  ) +
  theme(
    legend.position = "none",
    axis.text.x = element_text(angle = 25, hjust = 1) 
    )
```

## Nominal Statistics

### Summary of nominal variables (top categories)

```{r}
#| label: "Nominal Statistics"
#| tbl-cap: "Top categories for factor variables with counts, proportions, and mean price"
#| code-fold: true

top_n_per_var <- 10 

nominal_summary <- data |>
  clean_names() |>
  select(where(is.factor), price) |>
  pivot_longer(cols = where(is.factor),
               names_to = "Variable",
               values_to = "Category") |>
  group_by(Variable, Category) |>
  summarise(
    Count = n(),
    Percent = round(100 * Count / nrow(data), 2),
    Mean_Price = round(mean(price, na.rm = TRUE), 0),
    .groups = "drop"
  ) |>
  group_by(Variable) |>
  slice_max(order_by = Count, n = top_n_per_var, with_ties = FALSE) |>
  ungroup()

kable(
  nominal_summary,
  caption = paste0(
    "Top ", top_n_per_var,
    " categories per factor variable (counts, share %, and mean price)"
  ),
  digits = 2,
  align = c("l", "l", "r", "r", "r")
)
```

#### Interpretation
**city**: Most listings in Houston, Tucson, Phoenix, and Los Angeles. Prices range widely — highest in Los Angeles (~\$1.9M), lowest around \$250k (Saint Louis).

**state**: California, Texas, and Florida dominate listings (>30% total). California shows the highest mean price (~\$1.1M).

**status**: 55% for sale, 45% sold. Active listings are priced higher (~\$621k vs. \$515k).

**Overall**: Listings cluster in major U.S. cities and states, with strong regional price differences, especially high in California and large metro areas.

### Visualisation of nominal variables (top categories)

```{r}
#| label: "Visualization Nominal Statistics"
#| tbl-cap: "Top categories for factor variables with counts, proportions, and mean price"
#| code-fold: true

nominal_summary <- nominal_summary |>
  group_by(Variable) |>
  mutate(Category = forcats::fct_reorder(Category, Count),
         Category = factor(Category, levels = unique(Category))) |>
  ungroup()

# Plot: Facets untereinander, mit eigener x-Skala und y-Skala pro Variable
ggplot(nominal_summary, aes(x = Count, y = Category, fill = Variable)) +
  geom_col(show.legend = FALSE, alpha = 0.8, width = 0.7) +
  facet_wrap(~ Variable, ncol = 1, scales = "free", drop = TRUE) +
  scale_x_continuous(labels = label_comma()) +   #Tausendertrennung, keine 1e+05
  theme_minimal() +
  labs(
    title = "Top Categories per Factor Variable",
    x = "Count",
    y = "Category"
  ) +
  theme(
    panel.spacing.y = unit(1, "lines"),
    strip.text = element_text(size = 12, face = "bold"),
    axis.text.y = element_text(size = 8),
    plot.margin = margin(5, 15, 5, 5)
  )
```

# Bivariate Analysis

### Pairs Plot (all numeric variables)

```{r}
#| label: "Pairs Plot"
#| code-fold: true

set.seed(123)

data_num <- data |>
  janitor::clean_names() |>
  select(where(is.numeric)) |>
  slice_sample(n = 3000) |>
  mutate(across(everything(), log1p)) 

p <- ggpairs(
  data_num,
  progress = FALSE,
  upper = list(continuous = wrap("cor", size = 4, alignPercent = 0.8, stars = TRUE)),
  lower = list(continuous = wrap("points", alpha = 0.3, size = 0.7)),
  diag = list(continuous = wrap("densityDiag", alpha = 0.7))
)

p +
  theme_minimal(base_size = 11) +
  theme(
    strip.text = element_text(size = 8, face = "bold"),
    panel.grid = element_blank(),
    axis.text = element_text(size = 8),
    axis.title = element_text(size = 9),
    plot.title = element_text(face = "bold", size = 14, hjust = 0.5)
  ) +
  labs(title = "Pairs Plot (log-transformiert, n=3000)")
```

### Price vs. House Size by Status

```{r}
#| label: "scatter_price_size"
#| tbl-cap: "Price vs. house size by listing status"
#| code-fold: true

# Stichprobe ziehen für Performance
set.seed(123)
sample_data <- data %>% sample_n(50000)

plot_ly(
  sample_data,
  x = ~house_size,
  y = ~price,
  color = ~status,
  type = "scatter",
  mode = "markers",
  alpha = 0.6
) %>%
  plotly::layout(
    title = list(text = "Relationship Between House Size and Price by Status"),
    xaxis = list(title = "House Size (sqft)"),
    yaxis = list(title = "Price ($)", type = "log")
  )
```

#### Interpretation 

**Positive relationship**: Larger houses generally have **higher prices**, though the relationship weakens for very large properties.

**Status comparison**: Both for_sale and sold homes follow **similar trends**, but **for_sale** listings appear higher in price, suggesting sellers may list above sale values.

**High variation**: At similar sizes, prices vary widely — showing the strong influence of **location** and other factors.

**Outliers**: A few extremely large or expensive properties stretch the scale upward.


### Multiple Regression

```{r}
#| code-fold: true

model <- lm(price ~ house_size + bath + bed + state + status, data = data)
tidy(model) |>
  arrange(p.value) |>
  mutate(
    estimate = round(estimate, 1),
    std.error = round(std.error, 1),
    statistic = round(statistic, 1),
    p.value = signif(p.value, 3)
  ) |>
  datatable(
    caption = "Regressionsergebnisse (interaktiv)",
    filter = "top",         
    options = list(
      pageLength = 10,      
      autoWidth = TRUE,
      responsive = TRUE
    )
  )
```

```{r}
#| label: "Graphic Regression"
#| code-fold: true
 
tidy(model, conf.int = TRUE) |>
  filter(term != "(Intercept)") |>
  mutate(term = reorder(term, estimate)) |>
  ggplot(aes(x = estimate, y = term, fill = estimate > 0)) +
  geom_col(show.legend = FALSE) +
  geom_vline(xintercept = 0, linetype = "dashed") +
  theme_minimal() +
  labs(
    title = "Greatest contributors to price",
    x = "regression coefficient",
    y = ""
  )
```

### Average Property Price Map
```{r}
#| label: "state_mapping"
#| code-fold: true

valid_states <- tibble(
  state_name = c(state.name, "District of Columbia"),
  state_abbr = c(state.abb,  "DC")
)
```

```{r}
#| label: "map_avg_price"
#| tbl-cap: "Average property price by U.S. state"
#| code-fold: true

map_price <- data |>
  group_by(state) |>
  summarise(avg_price = mean(price, na.rm = TRUE), .groups = "drop") |>
  inner_join(valid_states, by = c("state" = "state_name")) |>
  mutate(avg_price_k = avg_price / 1000)

plot_ly(
  map_price,
  type = "choropleth",
  locationmode = "USA-states",
  locations = ~state_abbr,
  z = ~avg_price_k,
  text = ~paste0(state, "<br>Avg Price: $", round(avg_price_k, 1), "K"),
  colorscale = list(c(0, 1), c("lightblue", "darkblue")),
  colorbar = list(title = "Avg Price ($K)")
) |>
  plotly::layout(
    title = list(text = "Average Property Price by U.S. State"),
    geo = list(scope = "usa", projection = list(type = "albers usa"))
  )

```

#### Interpretation 

**Regional variation**: Western and coastal states show generally **higher property prices**, while central regions are lower.

**Highest averages**: States like **California, New York, and Washington** stand out with mean prices well above **$1M**.

**Moderate prices**: States such as Texas, Florida, and Arizona fall in the **mid-range** (~$400–650K).

**Lower averages**: Midwest and Southern states have more affordable properties on average.

**Summary**: Property values are heavily influenced by **geography** — with the highest prices concentrated along the coasts and major urban centers.

### Average House Size Map

```{r}
#| label: "map_avg_size"
#| tbl-cap: "Average house size by U.S. state"
#| code-fold: true

map_size <- data %>%
  group_by(state) %>%
  summarise(avg_size = mean(house_size, na.rm = TRUE), .groups = "drop") %>%
  inner_join(valid_states, by = c("state" = "state_name"))

plot_ly(
  map_size,
  type = "choropleth",
  locationmode = "USA-states",
  locations = ~state_abbr,
  z = ~avg_size,
  text = ~paste0(state, "<br>Avg Size: ", round(avg_size), " sqft"),
  colorscale = list(c(0, 1), c("lightgreen", "darkgreen")),
  colorbar = list(title = "Avg Size (sqft)")
) %>%
  plotly::layout(
    title = list(text = "Average House Size by U.S. State"),
    geo = list(scope = "usa", projection = list(type = "albers usa"))
  )
```

#### Interpretation 

**General trend**: Average house sizes are fairly consistent across most states, typically around 2,000–2,500 sqft.

**Larger homes**: Some central and mountain states (e.g., Colorado, Utah, Iowa) show slightly larger averages, possibly due to more available land.

**Smaller homes**: Coastal and densely populated states (e.g., New York, California) tend to have smaller average house sizes.

### Price Range by US State

```{r}
#| label: "map_extreme_price"
#| tbl-cap: "Price range (max − min) by U.S. state"
#| code-fold: true

map_extremes <- data |>
  group_by(state) |>
  summarise(
    min_price = suppressWarnings(min(price, na.rm = TRUE)),
    max_price = suppressWarnings(max(price, na.rm = TRUE)),
    .groups = "drop"
  ) |>
  mutate(range_price = max_price - min_price) |>
  inner_join(valid_states, by = c("state" = "state_name"))

plot_ly(
  map_extremes,
  type = "choropleth",
  locationmode = "USA-states",
  locations = ~state_abbr,
  z = ~range_price,
  text = ~paste0(
    state,
    "<br>Min: $", formatC(min_price, big.mark = ",", format = "f", digits = 0),
    "<br>Max: $", formatC(max_price, big.mark = ",", format = "f", digits = 0)
  ),
  colorscale = "Reds",
  colorbar = list(title = "Price Range ($)")
) |>
  plotly::layout(
    title = list(text = "Price Extrem Values by U.S. State (Max − Min)"),
    geo = list(scope = "usa", projection = list(type = "albers usa"))
  )
```

#### Interpretation 

**Highest ranges**: California shows by far the **largest price range** (over $400M), driven by extremely high luxury property values.

**Moderate ranges**: States like Florida and parts of the Northeast also show wide **price spreads**, reflecting diverse markets from affordable to luxury homes.

**Lower ranges**: Most central and midwestern states have smaller price gaps, indicating more uniform housing markets.

# Summary

